---
title: "R Notebook"
author: "Hamza Rashid"
email : "bolddusk@gmail.com"
description : "This notebook is for learning"
output: html_notebook
---

# Looping on the Command Line

Writing for, while loops is useful when programming but not particularly easy 
when working interactively on command line for data sets. There are some 
functions which implement looping to make life easier.
 - lapply: Loop over a list and evaluate a function on each element
 - sapply: Same as lapply but try to simplify the result.
 - apply: apply a function over the margin of an array.
 - tapply: apply a function over subsets of a vector.
 - mapply: Multivariate version of lapply.
 - An auxiliary function 'split' is also useful, particularly in conjunction 
    with 'lapply'.
    
lapply
 Apply loop on a list(lapply)
 lapply takes three arguments
 1- a list X;
 2- a function (or the name of the function) FUN;
 3- other arguments via its ... argument.
 
 If X is not a list, it will coerced to a list using as.list
 The acual looping is done internally in C code.
```{r}
a <- list(a=1:5, b=rnorm(10))
a
print("applying lapply with mean")
lapply(a,mean)

print("Seprately calculating mean")
mean(a$a)

# new list
b = list(a=1:4,b = rnorm(10),c=rnorm(20,1),d=rnorm(100,5))
print("List B")
lapply(b,mean)
```
 
```{r}
# random unified runif
a <- 1:5
lapply(a, runif)
```
 
# Topic 13.2 - Loop Sappy

sapply will try to simplify the result of lapply if possible.
 - if the result is list where every element is length 1, then a vector is 
    returned.
 - if the result is list where every element is a vector of the same length (>1)
    ,a matrix is returned.
 - if it can't figure things out, a list is returned.
```{r}
a <- list(a= 1:4, b = rnorm(10), c= rnorm(20,1), d= rnorm(100,5))

result_mean = sapply(a,mean)
result_mean


result_sum = sapply(a,sum)
print("result_sum")
result_sum
```
# Topic 13.3 - Loop appy

apply is used to evaluate function (often an anonymous one) over the margins
of an array
 - It is most often used to apply a function to the rows or columns of a matrix.
 - It can be used with general arrays, e.g. taking an average of an array of 
    matrices.
 - It is not really  faster than writing a loop, but it works in one line!
 
str(apply)
function(X,MARGIN, FUN,...)
 - X is an array.
 - MARGIN is an integer vector indicating which margins should be "retained". 
  - Only Rows & Columns
 - FUN is function to be applied.
 - ... is for other arguments to be passed to FUN.
 
```{r}
a <- matrix(rnorm(100,20,10))
result_mean = apply(a, 2,mean)
result_mean


result_sum = apply(a,2,sum)
print("sum result")
result_sum

```
 
 col/row sum and mean
 For sums and means of matrix dimensions, we have some shortcuts
  - rowSums = apply(x, 1, sum)
  - rowMeans = apply(x, 1, mean)
  - colSums = apply(x, 2, sum)
  - colMeans = apply(x, 2, mean)
```{r}
a <- matrix(1:10,5,2)
a
print("rowSums")
rowSums(a)
print("colSums")
colSums(a)

b <- matrix(rnorm(200),20,10)
print("Other way to use apply ")
apply(b, 1, quantile, probs = c(0.25,0.75) )
```
 
 
# Topic 13.4 - tapply
  tapply is used to apply a function over the subsets of a vector
  str(tapply)
  function(X, INDEX, FUN = NULL, ..., simplify = TRUE)
  - X is a vector.
  - INDEX is a factor or a list of factors (or else they are coerced to factors)
  - FUN is a function to be applied.
  - ... contains other arguments be passed to FUN
```{r}
a <- c(rnorm(10), runif(10), rnorm(10,1))
a
b <- gl(3, 10)
b

tapply(a,b,mean)
```

 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
