---
title: "R Notebook"
author: "Hamza Rashid"
email : "bolddusk@gmail.com"
description : "This notebook is for learning"
output: html_notebook
---

# Dates and Times in R

R has developed special representation of dates and times;
 - Dates are represented by the Date class.
 - Times are represent by the POSIXct or the POSIXlt class.
 - Dates are stored internally as the number of days since 1970-01-01.
 - Times are stored internally as the number of seconds since 1970-01-01.
 
 
Dates in R

Dates are represented by the Date class and can be coerced by from a character
string using th as.Date() function.
```{r}
a = as.Date("2021-01-01")
a

b = as.POSIXct("1970-01-01")
b

c = as.POSIXlt("1970-01-01")
c
```
Times are represented using the POSIXct or the POSIXclt class.
  - POSIXct is just a very large integer under the hood; it is useful class when
    class when you want to store times in something like data frame.
  - POSIXlt is a list underneath and it stores a bunch of other useful 
    information like day of the week, day of the year,month, day of the month
    etc.
  
There are few generic functions what work on Date on Times.
 - weedays: give the day of the week.
 - months: give the month name.
 - quarters: give the quarter number ("Q1","Q2","Q3","Q4").
 
 
 Times can coerced from a character string using the as.POSIXlt or POSIXct
 funtion.
```{r}
a <- Sys.time()
a

psx = as.POSIXlt(a)
psx

names(unclass(psx))

psx$sec
psx$wday

pct = as.POSIXct(a)
pct
```
 
There is 'strptime()' function in case your dates are written in a different format.
- ?strptime if you don't remember formatting string
```{r}
datestring <- c("January 10, 2021 10:40","December 10, 2021 11:40")
a <- strptime(datestring, "%B %d, %Y %H:%M")
a

```
Operations on Dates and Times
Mathematical operations can be used on Dates & Times.
```{r}
a <- as.Date("2021-05-06")
b <- strptime("10 Jan 2021 11:10:20", "%d %b %Y %H:%M:%S")
#a-b
class(a)
class(b)
a <- as.POSIXct(a)
a-b
```
```{r}
a <- as.Date("2012-03-01")
b <- as.Date("2012-04-25")
a-b

# Example of Time difference
c <- as.POSIXct("2012-10-25 10:00:00")
d <- as.POSIXct("2012-10-25 16:20:00")
d-c
```


























