---
title: "R Notebook"
author: "Hamza Rashid"
email : "bolddusk@gmail.com"
description : "This notebook is for learning"
output: html_notebook
---
# Control Structures
# Topic 8.1 - If Else

Control structures in R allow you to control the flow of execution of the program, depending on runtime conditions.
Common structures are:

 - if, else: testing a condition
 - for: execute a loop fixed number of times
 - while: execute a loop while a condition is true.
 - repeat: execute an infinite loop.
 - break: break the execution of loop.
 - next: skip an interaction o loop.
 - return: exit a function.
 
- Most control structures are not used in interactive sessions, but rather when writing functions or longer expressions.

```{r}
a <- 4
b <- 5
a
b
if(a > 4){
  b = 8
}else{
  b = -4
}
a
b
c <- if(a > 4){ 100 }else{-100}
c
```
# Topic 8.2 - For Loop

for loops take an iterative variable and assign it successive values from a sequence or vector.for loops are most used for iterating over the elements of an object(list, vector, etc.).
```{r}
for(i in 1:5){
  print(i)
}

a = c("a","b","c","d","e")
for(i in a){
  print(i)
}

print("using seq_along")
for( i in seq_along(a)){
  print(a[i])
}

print("single line loop")
for(i in 1:4) print(a[i])
```
Nested Loop
```{r}
a <- matrix(1:12,3,4)
a
print("Printing a row wise")
for(i in seq_len(nrow(a))){
  for(j in seq_len(ncol(a))){
    print(a[i,j])
  }
}
```

# Topic 8.3 - While Loop

While loops begin by testing the condition. If it is true, then they execute the loop body.Once the loop is executed, the condition is tested again, and so forth.
```{r}
count <- 0
while(count < 5){
  print(count)
  count = count+1
}
```
```{r}
a <- 5
while (a >=3 && a <=8){
  print(a)
  coin <- rbinom(1,1,0.5)
  if(coin == 1){
    a <- a+1
  }else{
    a <- a-1
  }
}
```
 










