---
title: "R Notebook"
author: "Hamza Rashid"
email : "bolddusk@gmail.com"
description : "This notebook is for learning"
output: html_notebook
---
# Control Structures
# Topic 8.1 - If Else

Control structures in R allow you to control the flow of execution of the program, depending on runtime conditions.
Common structures are:

 - if, else: testing a condition
 - for: execute a loop fixed number of times
 - while: execute a loop while a condition is true.
 - repeat: execute an infinite loop.
 - break: break the execution of loop.
 - next: skip an interaction o loop.
 - return: exit a function.
 
- Most control structures are not used in interactive sessions, but rather when writing functions or longer expressions.

```{r}
a <- 4
b <- 5
a
b
if(a > 4){
  b = 8
}else{
  b = -4
}
a
b
c <- if(a > 4){ 100 }else{-100}
c
```
# Topic 8.2 - For Loop

for loops take an iterative variable and assign it successive values from a sequence or vector.for loops are most used for iterating over the elements of an object(list, vector, etc.).
```{r}
for(i in 1:5){
  print(i)
}

a = c("a","b","c","d","e")
for(i in a){
  print(i)
}

print("using seq_along")
for( i in seq_along(a)){
  print(a[i])
}

print("single line loop")
for(i in 1:4) print(a[i])
```
Nested Loop
```{r}
a <- matrix(1:12,3,4)
a
print("Printing a row wise")
for(i in seq_len(nrow(a))){
  for(j in seq_len(ncol(a))){
    print(a[i,j])
  }
}
```

# Topic 8.3 - While Loop












